### 2022-12-06 README.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2

>>>>

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
depth2 = depth^2/depth[1]^2 #normalize for stability
d18O = NGRIP_5cm$d18O

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,d18O=d18O)
formula = dy~-1+depth2+d18O

### 2022-12-06 inst/Examples/Paper_results/Paper_results_1.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r bremla, eval=TRUE}
library(bremla)
library(dplyr)
library(ggplot2)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
n=length(diff(age))
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

>>>>

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r bremla, eval=TRUE}
library(bremla)
library(dplyr)
library(ggplot2)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
depth2 = depth^2/depth[1]^2 #normalize for stability
proxy = NGRIP_5cm$d18O
n=length(diff(age))
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

### 2022-12-06 inst/Examples/Paper_results/Paper_results_2.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r initialize, eval=FALSE}
#library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
n=length(diff(age))
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)
# plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
     # xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
formula = dy~-1+depth2+proxy
```

@@

```{r simex}
rm(list=ls())
require(stats)
library(ggplot2)
set.seed(123)
n=1000
y0 = 1000
z0 = 1200
phi=0.8; sigma=1.2
depth = z0+seq(from=0.5,500,length.out=1000)
dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
events=list(locations=c(1,250,750,1000))
a1 = c(-rep(1,249)*15,numeric(751))
a2 = c(numeric(249),rep(1,500)*50,numeric(251))
a3 = c(numeric(749),-rep(1,251)*110)
c1 = c(depth[1:249]/60,numeric(751))
c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
c3 = c(numeric(749),depth[1:251]/10)
dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth^2); data=rbind(c(y0,z0,0,0,0),df)
formula=dy~-1+proxy#+depth2
nsims = 10000
synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))
res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             events=list(locations=c(1,250,750,1000),locations_unit="index"),
             control.fit=list(noise="ar1"),
             synchronization=synchronization,
             control.sim = list(synchronized=TRUE,
                                nsims=nsims))
par(mfrow=c(1,2))
plot(res$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(sigma),ylab="Density",main=expression(paste("(a) Posterior distribution of ",sigma)))
abline(v=res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
           res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),lwd=1.5,col="red")
plot(res$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density",main=expression(paste("(b) Posterior distribution of ",phi)))
abline(v=res$fitting$inla$hyperparameters$results$phi$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$phi$quant0.025,
           res$fitting$inla$hyperparameters$results$phi$quant0.975),lwd=1.5,col="red")
par(mfrow=c(1,1))
#plot results

@@

## Real data example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)
results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=list(method="adolphi"),
                control.fit=list(method="inla"),
                control.sim=list(synchronized=TRUE) )
summary(results)
```

>>>>

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r initialize, eval=FALSE}
#library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
depth2 = depth^2/depth[1]^2 #normalize for stability
proxy = NGRIP_5cm$d18O
n=length(diff(age))
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)
# plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
     # xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
formula = dy~-1+depth2+proxy
```

@@

```{r simex}
rm(list=ls())
require(stats)
library(ggplot2)
set.seed(123)
n=1000
y0 = 1000
z0 = 1200
phi=0.8; sigma=1.2
depth = z0+seq(from=0.5,500,length.out=1000)
depth2 = depth^2/depth[1]^2 #normalize for stability

dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
events=list(locations=c(1,250,750,1000))
a1 = c(-rep(1,249)*15,numeric(751))
a2 = c(numeric(249),rep(1,500)*50,numeric(251))
a3 = c(numeric(749),-rep(1,251)*110)
c1 = c(depth[1:249]/60,numeric(751))
c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
c3 = c(numeric(749),depth[1:251]/10)
dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth2); data=rbind(c(y0,z0,0,0,0),df)
formula=dy~-1+proxy#+depth2
nsims = 10000
synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))
res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             events=list(locations=c(1,250,750,1000),locations_unit="index"),
             control.fit=list(noise="ar1"),
             synchronization=synchronization,
             control.sim = list(synchronized=TRUE,
                                nsims=nsims))
par(mfrow=c(1,2))
plot(res$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(sigma),ylab="Density",main=expression(paste("(a) Posterior distribution of ",sigma)))
abline(v=res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
           res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),lwd=1.5,col="red")
plot(res$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density",main=expression(paste("(b) Posterior distribution of ",phi)))
abline(v=res$fitting$inla$hyperparameters$results$phi$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$phi$quant0.025,
           res$fitting$inla$hyperparameters$results$phi$quant0.975),lwd=1.5,col="red")
par(mfrow=c(1,1))
#plot results

@@

## Real data example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")
age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
depth2 = depth^2/depth[1]^2 #normalize for stability
proxy = NGRIP_5cm$d18O
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
formula = dy~-1+depth2
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)
results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=list(method="adolphi"),
                control.fit=list(method="inla"),
                control.sim=list(synchronized=TRUE) )
summary(results)
```
